---
title: "Getting data ready for annotation"
author: "Bodo"
date: "04/08/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This script loads in the combined data and prepares it for annotation by doing the following steps:

- add unique ID
- randomize
- add columns for annotation

## Load packages and data

Load packages:

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
```

Load data:

```{r, warning = FALSE, message = FALSE}
birds <- read_csv('../data/voice_all_guides.csv')
```

## Processing

Add unique id:

```{r}
birds$unique_id <- str_c('id_', 1:nrow(birds))
```

Re-order columns and exclude those that irrelevant for annotation (length and species info):

```{r}
birds <- select(birds,
                unique_id, fieldguide, english_common, voice)
```

Randomize:

```{r}
set.seed(42)
birds <- sample_n(birds, size = nrow(birds))
```

Replace '…' which is not displayed correctly in Excel with three separate dots:

```{r}
birds <- mutate(birds,
                voice = str_replace_all(voice, '…', '...'))
```

Add relevant columns for annotation. First, a column which is a copy of the 'voice' section where boundaries (semicolons) will be placed in between different types of sound objects.

```{r}
birds$voice_bounded <- birds$voice
```

Second, a column coding for whether a typo has been fixed. Crucially, if there is a typo, this needs to be fixed in both `voice` and `voice_bounded`:

```{r}
birds$typo <- 'NO'
```

Third, an empty column to flag potential cases of Warblish we have missed:

```{r}
birds$warblish <- ''
```

Fourth, an empty column telling us if there's any interesting observations or anything specific about how this example has been processed, let's call it `comment`:

```{r}
birds$comment <- ''
```

Save this:

```{r}
write_csv(birds, '../data/voice_all_guides_for_annotation.csv')
```

This completes this script.





