---
title: "Combine all field guides"
author: "Bodo"
date: "21/07/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This script takes the folder `raw_data` in the `data` folder, loops through all file names and combines them into one big table.

## Processing

Load package:

```{r message = FALSE, warning = FALSE}
library(tidyverse)
```

Get file names from `data/raw_data` folder:

```{r}
filenames <- list.files('../data/raw_data')
```

Loop through and get all files. The loop has to branch out because not all field guides have length info.

```{r}
# Create empty data frame to be filled:

df <- c()

# Loop through file names:

for (i in seq_along(filenames)) {
  
  x <- read_csv(str_c('../data/raw_data/', filenames[i]))
  
  if ('Length' %in% colnames(x)) {
    x <- select(x, EnglishCommon, LatinSpecies, Voice, Length)
  } else {
    x <- select(x, EnglishCommon, LatinSpecies, Voice)
    x$Length <- NA
  }
  
  # Add fieldguide info from file name:
  
  x$fieldguide <- filenames[i]
  
  # Combine with growing data frame:
  
  df <- rbind(df, x)
}
```

Rename and re-order columns:

```{r}
df <- df %>% 
  rename(english_common = EnglishCommon,
         latin_species = LatinSpecies,
         voice = Voice,
         length = Length) %>% 
  select(fieldguide, english_common, latin_species, length, voice)
```

Get rid of the `.csv` from the `fieldguide` column:

```{r}
df <- mutate(df,
             fieldguide = str_replace_all(fieldguide, '\\.csv', ''))
```

Write externally:

```{r}
write_csv(df, '../data/voice_all_guides.csv')
```




