---
title: "Merging field guides with IOC world list"
author: "Bodo"
date: "21/07/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this script, I will merge data from our 18 field guides with the IOC world bird list. Unfortunately the IOC list is quite a messy data frame - not particularly machine-readable due to empty rows etc. So some data cleaning will be necessary.

## Setup

Load packages:

```{r warning = FALSE, message = FALSE}
library(tidyverse)
```

Load data:

```{r warning = FALSE, message = FALSE}
# Field guide data:

voice <- read_csv('../data/voice_all_guides.csv')

# IOC word list:

IOC <- read_csv('../data/master_ioc_list_v12.1.csv',
                skip = 3) # first two rows are empty
```

## Data wrangling IOC word list

Check the structure of this data frame:

```{r}
IOC
```

Let's rename columns. First make all lowercase, then get rid of those pesky brackets.

```{r}
# All lowercase:

colnames(IOC) <- str_to_lower(colnames(IOC))

# Get rid of brackets:

colnames(IOC) <- str_replace_all(colnames(IOC),
                                 '(\\()|(\\))', # regex for all brackets
                                 '') # replace with nothing

# Replace spaces with underscores:

colnames(IOC) <- str_replace_all(colnames(IOC),
                                 ' ', '_')

# Make some of the names shorter:

IOC <- rename(IOC,
              family = family_scientific,
              english = species_english)
```

Next, the `NA`'s need to be filled downwards.

```{r}
IOC <- fill(IOC,
            order, genus, species_scientific, subspecies,
            family, family_english,
            english)

# Check:

IOC
```

Make the `order` column nicer (right now it's all caps, which screams at me!):

```{r}
IOC <- mutate(IOC,
              order = str_to_title(order))
```


To create a species identifier that can be matched with our field guide database, concatenate the strings in the `genus` column together with the strings in the `species_scientific` column with a space in between:

```{r}
IOC <- mutate(IOC,
              latin_species = str_c(genus, ' ', species_scientific))

# Check 10random rows:

sample_n(IOC, 10) %>% select(latin_species,
                             order, family,
                             english)
```

This looks usable.

Let's reduce this `IOC` tibble to only those columns that we want to appear in the cleaned write-out. Then save this externally:

```{r}
# Select columns:

IOC <- select(IOC,
              latin_species, order, family, family_english, english, subspecies)

# Write:

write_csv(IOC, '../data/master_ioc_list_v12.1_cleaned.csv')
```

An issue with merging the data is going to be the fact that each subspecies is on a different row. As nearly all field guides have no subspecies information, we will dispense of this and get rid of duplicates. After that, we can drop the `subspecies` column:

```{r}
IOC <- filter(IOC, !duplicated(latin_species)) %>% 
  select(-subspecies)
```

Check:

```{r}
IOC
```

The row of NAs is not an issue.

## Merge field guide data with IOC world list

Merge the new `IOC` tibble with the field guide:

```{r}
voice <- left_join(voice, IOC, # both tibbles
                   by = 'latin_species') # key
```

## Check NAs and append coding columns:

How many did not merge?

```{r}
filter(voice, is.na(family)) %>% nrow() # quite a few!

filter(voice, is.na(family))
```

Quite a few did not merge correctly.

Let's flag these by creating an identifier variable:

```{r}
voice <- mutate(voice,
                match = complete.cases(family))
```

Append columns for what Nigel should add. First, a copy of the `latin_species` column that will contain the corrected name:

```{r}
voice <- mutate(voice,
                latin_species_correct = latin_species)
```

Create a column for whether a name was changed (default populated with `no`), then a column where comments can be added:

```{r}
voice$changed <- 'no'
voice$comment <- ''
```

Write this to file:

```{r}
write_csv(voice, '../data/voice_all_guides_with_ioc.csv')
```

Another approach of data cleaning in bulk would be to count all latin species names and sort them alphabetically. This way we can easily spot typos:

```{r}
species_counts <- voice %>%
  count(latin_species) %>% arrange(latin_species)

# Save outside:

write_csv(species_counts, '../data/all_guides_species_counts.csv')
```






